Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-08-16 paper/letter.tex
title, authors, affiliations and exact values

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Þú ert jörðin}
\\
\\
JEFR$^{1}$, MG$^3$, MCRG$^1$, JNGC$^2$, MD$^3$
\\
\\
\scriptsize
{$^1$ Bogota
\\
$^2$ Tucson
\\
$^3$ Oslo
\normalsize
\\
\\
\textbf{
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad symmetric Lyman-$\alpha$ emission that had evaded theoretical
interpretation so far. 
In this letter we explain these features by two different models: an
homogeneous gaseous sphere undergoing bulk rotation and an interstellar
medium composed by clumps with random motions.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these physical conditions.
We find that both models independently require high velocities that
translate into a dynamical mass at least $20$ times larger than the
neutral mass hydrogen in the galaxy. 
We argue that a plaussible explanation for the high dynamical mass is
the presence of a super-massive black hole.  
This work demonstrates the importance of considering rotation and
multiphase physics under the possible conditions shaping the
\lya\ spectra of the first galaxies. 
Additionally, if future kinematic maps of \tol\ confirm the high
velocities postulated in our model, it would provide new
evidence for dwarf galaxies as hosts of supermassive black
holes.  
}

@@

\tol\ is a compact star forming dwarf galaxy that presents a
strong Ly-$\alpha$ emission \cite{Thuan97} with two puzzling 
features: it is symmetric and single peaked.
Commonly, the Ly-$\alpha$ line has a single or asymmetric double peak. 
These two special features in \tol\ had evaded a physical
interpretation so far. Using either bulk rotation or a multiphase ISM,
we manage to explain the \lya\ line in \tol.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{CLARA-TOL-main.pdf}
\caption{{\bf Broadn, single peaked and symmetric Ly-$\alpha$ emission of \tol.}
  Dots correspond to the observational data. The line shows the results
of our best model from a full radiative transfer simulation both for the rotation and multiphase models.}
\end{center}
\end{figure}

@@

The analytical solution for the rotation sphere was the base to
perform the Markov Chain Monte Carlo Calculation using the
\texttt{emcee} implementation.
We explore flat priors on $V_{\rm max}$ $\log_{10}\tau$, $\log_{10} T$ and
$\theta$ using $XX$ steps.
The results are summarized in Figure $XX$. 
From this model we find that the fiducial parameters that could
explain the broad features in \tol\ are $V_{\rm max}$, $\log \tau
=XX$, $\log_{10} T = $ and  $\theta =$

>>>>

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Lyman-$\alpha$ sketches a massive dwarf}
\\
\\
Jaime E. Forero-Romero$^{1}$, Max Groenke$^2$, Mar\'ia Camila
Remolina-Guti\'errez$^1$, Juan Nicol\'as
Garavito-Carmargo$^3$, Mark Dijkstra$^2$.
\\
\\
\scriptsize
{$^1$ Departamento de Física, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia 
\\
$^2$ Institute of Theoretical Astrophysics, University of Oslo,
Postboks 1029 Blindern, NO-0315 Oslo, Norway.
\\
$^3$ Department of Astronomy, University of Arizona, 933 North Cherry
Avenue, Tucson, AZ 85721, USA. 
\normalsize
\\
\\
\textbf{
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad symmetric Lyman-$\alpha$ emission that had evaded theoretical
interpretation so far. 
In this letter we explain these features by two different models: an
homogeneous gaseous sphere undergoing bulk rotation and an interstellar
medium composed by outflowing clumps with additional random motions.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these physical conditions.
We find that both models independently require high velocities
(either a bulk rotation of $348^{+75}_{-48}$ km s$^{-1}$ or a clump velocity
dispersion of $71^{+17}_{-25}$ km s$^{-1}$ with outflows of
$79^{+167}_{-60}$ km s$^{-1}$) consistent with a dynamical mass on
the order of $5\times 10^9$ solar masses, which is at least $20$ times
larger than the neutral mass hydrogen in the galaxy.   
We argue that a plaussible explanation for the high dynamical mass is
the presence of a super-massive black hole.  
This work demonstrates the importance of considering rotation and
multiphase physics among the possible conditions shaping the
\lya\ spectra of the first galaxies. 
Additionally, if future kinematic maps of \tol\ confirm the high
velocities postulated in our model, it would provide new
evidence for dwarf galaxies as hosts of supermassive black
holes.  
}

@@

\tol\ is a compact star forming dwarf galaxy that presents a
strong Ly-$\alpha$ emission \cite{Thuan97} with two puzzling 
features: it is symmetric and single peaked.
Commonly, the Ly-$\alpha$ line has a single or asymmetric double peak. 
These two special features in \tol\ had evaded a physical
interpretation so far. 
Including the effects of either bulk rotation or a multiphase ISM, we
manage to explain the \lya\ line in \tol.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{CLARA-TOL-main.pdf}
\caption{{\bf Broad, single peaked and symmetric Ly-$\alpha$ emission of \tol.}
  Dots correspond to the observational data. The line shows the results
of our best model from a full radiative transfer simulation both for the rotation and multiphase models.}
\end{center}
\end{figure}

@@

The analytical solution for the rotation sphere was the base to
perform the Markov Chain Monte Carlo Calculation using the
\texttt{emcee} implementation.
We explore flat priors on $V_{\rm max}$ $\log_{10}\tau$, $\log_{10} T$ and
$\theta$ using $XX$ steps.
The results are summarized in Figure $XX$. 
From this model we find that the fiducial parameters that could
explain the broad features in \tol\ are $V_{\rm max}=348^{+75}_{-48}$
km s$^{-1}$, $\log \tau = 6.96^{+0.26}_{-0.18}$, $\log_{10} T/\mathrm
{K} = 4.27^{+0.11}_{-0.18}$ and  $\theta = 35.78^{+2.13}_{-1.88}$ degrees.

\begin{figure}
\begin{center}
\includegraphics[width=1.0\textwidth]{emcee_results.pdf}
\caption{{\bf Results from the Markov Chain Monte Carlo compuation for the rotation model.}}
\end{center}
\end{figure}
